Full-day, in home validation of infant body position measurements from inertial sensors

### 2023-03-14 manuscript.Rmd
revised everything

The same machine learning classification process was used as in prior work [@FranchakScott2021]. Using the synchronization point, human-coded body annotations from video were linked to the corresponding times in the IMU timeseries data. A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code.

@@

Windowed datasets were used for machine learning classification and validation. For each analysis reported in the results, a subset of data were defined as a "training" set and another, independent portion of the data were defined as a "testing" set. Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set. Random forest models [@Breiman2001] were trained to predict the human coded body position label for each window from the set of 436 motion features using the *randomForest* package [@LiawWiener2002] in R version 4.1.2 [@R42]. Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit. The *tidymodels* set of packages in R [@tidymodels] was used to tune the "mtry" and "ntrees" hyperparameters, which control the number of columns sampled and the total number of trees created when modeling. We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance. Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the *predict* function. Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.

@@

We report four sets of results based on `r length(unique(ppts$id_uni))` full-day testing sessions resulting in a total of `r sum(duration$duration_hours)` hours of movement recording. First, we focus on the accuracy of group and individual models trained and validated on the proximal comparison period. This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events. Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available. This distal comparison is crucial, because it provides the first ever test of how well body classification models predict behavior beyond the initial, supervised period. Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day should be valid. In the third set of results we examine the data quality of full-day recordings, and in the fourth set of results we test whether full-day recordings capture well-established age differences in body position experiences.

@@

metric_plot <- function(ds, met) {
  ds %>% filter(metric == met) %>% 
    ggplot(aes(x = Model, y = value)) + 
    geom_sina(alpha = .7, fill = "turquoise3", size = 4, maxwidth = .5, shape = 21) +
    stat_summary(fun = "mean", mapping = aes(x = Model, y = value), shape = "-", fill = "black", size = 4) + 
    ylim(0, 1) + ylab(met) + xlab("") + 
    theme(plot.title.position = "plot",
          axis.title.y = element_text(size = 13),
          axis.text.y = element_text(size = 11),
          axis.text.x = element_text(size = 13))
}
# metric_labels <- c("Overall Accuracy", "Balanced Accuracy", "F1", "Sensitivity", "Pos Pred Value", "Kappa")
# title_labels <- c("A. Overall Accuracy", "B. Balanced Accuracy", "C. F1 Score", "D. Sensitivity", "E. Pos. Pred. Value", "F. Kappa")
metric_labels <- c("Overall Accuracy", "Kappa", "Sensitivity", "Pos Pred Value")
title_labels <- c("A. Overall Accuracy", "B. Kappa", "C. Sensitivity", "D. Pos. Pred. Value")
metric_plots <- map(metric_labels, ~ metric_plot(metrics_long, .x)) %>% map2(., title_labels, ~.x + ggtitle(.y))
(metric_plots[[1]] + metric_plots[[2]])
# (metric_plots[[1]] + metric_plots[[2]])/(metric_plots[[3]]+metric_plots[[4]])/(metric_plots[[5]] + metric_plots[[6]])

@@

The first set of analyses use data from the proximal comparison, in which video annotations and motion data were tightly synchronized (within 30-60 ms). This high degree of synchronization makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work [@FranchakScott2021], we compared two types of models: *group models* and *individual models*. To assess the accuracy of body position classification for each recording session, we held back the last 25% of a session's proximal comparison data as the testing set. The testing set was never used as training data. A group model was trained for each session using an aggregated dataset containing the first 75% of all **other** sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set. In contrast, individual models were trained using the first 75% of a recording session, then tested against the last 25%. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches. Figure \@ref(fig:metrics), Table \@ref(tab:metrics-table), and Table \@ref(tab:metrics-by-class) summarize the performance of group and individual models using standard metrics for classification.

@@

Beyond accuracy, it is important to establish that classification is neither too sensitive nor too conservative in choosing one position versus another. Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?). High sensitivity means that events are unlikely to be missed. In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded sitting events?). High PPV means that we can be confident in the event label. Table \@ref(tab:metrics-by-class) shows the sensitivity and PPV by body position class for group and individual models. For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held. All values were sufficient (> .8). Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held. Overall, the results suggest a reasonable balance between sensitivity and PPV among different body position classes for both model types. Table \@ref(tab:metrics-table) shows the overall sensitivity and PPV across classes.

@@

In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording. In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period---the 90-minute long recordings that followed the proximal recordings (Figure \@ref(fig:exemplar-timeline)). A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used. As previously mentioned, distal videos had only coarse temporal synchrony with motion recordings (misalignment of up to ~ 5 s in either direction). This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work [@YaoPlotz2019; @FranchakScott2021]. For each distal comparison, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding. Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.

Whereas the first set of analyses used all `r length(unique(ppts$id_uni))` sessions, this was not possible in the distal comparison. Because the start of the visit was scheduled during a time the infant was awake, it was common for a nap to be needed following the proximal period. Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording. Three additional sessions were excluded because a caregiver accidentally turned off the video camera (*n* = 1) or left the house (*n* = 2). This left `r length(unique(agree$group$file))` sessions with usable distal comparison data.

@@

Visual inspection of Figure \@ref(fig:part2overall) indicated two extreme outliers, which we marked by a gray square and a gray diamond. We closely investigated each of the two outliers to understand why their agreement was below the standard of the other sessions. The "gray square" outlier had significant confusion between sitting and supine classification. Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine. The infant also spent a long time in mother's arms in an ambiguous supine/sitting position that was hard to classify. This participant's proximal accuracy was also poor because similar ambiguous appeared during the initial recording. In contrast, the "gray diamond outlier" had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories. Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as "upright" but the models predicted it as "held". Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infant's moved while walking upright.

@@

```{r part2bins, warning=FALSE, fig.cap="Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison. Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins. Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute. Performance is plotted separately for (A) group models and (B) individual models. The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=8, fig.height=4.5, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
calculate_agreement_bins <- function(ds) {
  ds <- ds %>% group_by(file) %>% 
    mutate(sample = row_number(), bin = floor(sample/300)) %>% 
    add_count(file, bin) %>% ungroup() %>% filter(n > 200) %>% group_by(file) %>% mutate(num_bins = length(unique(bin))) %>% ungroup()
  model <- ds %>% count(file, bin, num_bins, pos) %>% add_count(file, bin,num_bins,  wt = n, name = "total") %>% mutate(prop_model = n/total)
  human <- ds %>% count(file, bin, num_bins,  code) %>% add_count(file, bin, num_bins, wt = n, name = "total") %>% mutate(prop_human = n/total)
  agree <- full_join(select(model, file, pos, prop_model, bin, num_bins), 
                     select(human, file, code, prop_human, bin, num_bins), 
                     by = c("file" = "file", "pos" = "code", "bin" = "bin", "num_bins" = "num_bins"))
  agree <- agree %>% mutate(prop_human = ifelse(is.na(prop_human), 0, prop_human),
                            prop_model = ifelse(is.na(prop_model), 0, prop_model),
                            pos - factor(pos, levels = pos_levels))
  agree <- complete(agree, nesting(file, bin, num_bins), pos, fill = list(prop_human = 0, prop_model = 0))
}

@@

Collecting real-time data that can be synchronized to other sources is a key goal in our development of long-form body position classification. Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale. We repeated agreement analysis, correlating human-coded body position prevalence with model-predicted prevalence, after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)). Infants had varying numbers of 10-minute bins depending on much time they were awake and on camera. Bins were included only if there was > X minutes of usable data.
<!-- RANGE OF BINSm USABLE DATA -->

Table \@ref(tab:pt2binstable) shows the agreement correlation coefficients for group and individual models, including and excluding the outliers identified in the previous section. Performance at a short timescale was similar to performance overall: Overall agreement after excluding outliers was excellent for both group (*r* = `r group_agree_cor_noout`) and individual models (*r* = `r split_agree_cor_noout`). Within-position correlations were weakest for held and strongest for upright regardless of the model type. Agreement for prone was better for group models, whereas sitting and supine were better predicted by individual models.

@@

Among participants who wore the sensors from morning to bedtime, the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \@ref(fig:timelines)). The total length of the recording period ranged from `r apa_num(min(coverage$total))`-`r apa_num(max(coverage$total))` hours with a median of `r apa_num(median(coverage$total))` hours. Nap times reported during the recording period ranged from `r apa_num(min(coverage$naptime))`-`r apa_num(max(coverage$naptime))` hours with a median of `r apa_num(median(coverage$naptime))` hours. Body position data was available to describe `r apa_num(min(coverage$prop_accounted))`-`r apa_num(max(coverage$prop_accounted))`% of the awake portion of the recording period (median = `r apa_num(median(coverage$prop_accounted))`%). Younger infant \#13 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home. Overall, the recordings produced a median of `r apa_num(median(coverage$hours_recorded))` hours of motion data, with the entire dataset totaling `r apa_num(sum(coverage$hours_recorded))` hours. Even the shortest full recording (`r apa_num(min(coverage$hours_recorded))` hours) exceeded the longest observational sessions in past work in which an experimenter operated a camera.

@@

The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically. In @Survey, upright time increased from 0.6-5.5% of the time for infants 3-6 months, increasing to 22.0% at 12 months. Similarly, younger infants' sitting time estimated from full-day group models was `r pos_sum_lookup["Upright", "Younger_Group"]`% compared with `r pos_sum_lookup["Upright", "Older_Group"]`% for the older group. In @Survey, prone time increased ranged from 2.9-9.2% of the time for infants 3-6 months, and was 7.2% at 12 months. In the current study, younger infants' prone time was was `r pos_sum_lookup["Prone", "Younger_Group"]`% and `r pos_sum_lookup["Prone", "Older_Group"]`% for older infants. Held, sitting, and supine are more difficult to compare because prior work [@Survey] considered held any time that infants were in caregivers' arms or on their laps, however, we only counted laying or sitting on caregivers' laps as supine and sitting, respectively. Further complicating the comparison, @Survey included a reclined category that we would classify as either supine or sitting depending on the specifics. Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants. Sitting was greater for older infants (`r pos_sum_lookup["Sitting", "Older_Group"]`%) compared with younger infants (`r pos_sum_lookup["Sitting", "Younger_Group"]`%). Moreover, sitting was the most frequent body position followed by upright in 12-month-olds [@Survey], just as we observed in the current study. Overall, the estimates from the current investigation are consistent with age differences found using a different methodology.

>>>>

The same machine learning classification process was used as in prior work [@FranchakScott2021]. Using the synchronization point, human-coded body annotations from video were linked to the corresponding times in the IMU timeseries data. A single, merged dataset was created with synchronized accelerometer signals (in three orientations: X, Y, and Z) and gyroscope signals (in three orientations: roll, pitch, and yaw) for each of the four sensors (left hip, right hip, left ankle, right ankle) with the corresponding timestamp and body position code using the *timetk* package [@DanchoVaughan2023] and the *lubridate* package [@GrolemundWickham2011] in R version 4.1.2 [@R42].

@@

Datasets containing one row per 4-s window were used for machine learning classification and validation. For each analysis reported in the results, a subset of data were defined as a "training" set and another, independent portion of the data were defined as a "testing" set. Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set. Random forest models [@Breiman2001] were trained to predict the human coded body position label for each window from the set of 436 motion features using the *randomForest* package [@LiawWiener2002]. Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit. The *tidymodels* set of packages in R [@tidymodels] was used to tune the "mtry" and "ntrees" hyperparameters, which control the number of columns sampled and the total number of trees created when modeling. We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance. Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the *predict* function. Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.

@@

We report three sets of results based on `r length(unique(ppts$id_uni))` full-day testing sessions resulting in a total of `r round(sum(duration$duration_hours))` hours of movement recording. First, we focus on the accuracy of group and individual models trained and validated from data during the proximal comparison period. This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events. Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available. This distal comparison is critical because it provides the first-ever test of how well infant body classification models predict behavior beyond the initial, supervised period. Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day will be valid. In the third set of results we examine the data quality of full-day recordings and test whether full-day recordings capture well-established age differences in body position experiences.

@@

metric_plot <- function(ds, met) {
  ds %>% filter(metric == met) %>% 
    ggplot(aes(x = Model, y = value)) + 
    geom_sina(alpha = .7, fill = "turquoise3", size = 4, maxwidth = .5, shape = 21) +
    stat_summary(fun = "mean", mapping = aes(x = Model, y = value), shape = "-", fill = "black", size = 4) + 
    ylim(0, 1) + ylab(met) + xlab("") + 
    theme(plot.title.position = "plot",
          axis.title.y = element_text(size = 13),
          axis.text.y = element_text(size = 11),
          axis.text.x = element_text(size = 13))
}

metric_labels <- c("Overall Accuracy", "Kappa", "Sensitivity", "Pos Pred Value")
title_labels <- c("A. Overall Accuracy", "B. Kappa", "C. Sensitivity", "D. Pos. Pred. Value")
metric_plots <- map(metric_labels, ~ metric_plot(metrics_long, .x)) %>% map2(., title_labels, ~.x + ggtitle(.y))
(metric_plots[[1]] + metric_plots[[2]])

@@

The first set of analyses use data from the proximal comparison to determine the "best case" accuracy of the models, training and testing on similar types of data. The high degree of temporal synchronization between video and motion data during this period makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work [@FranchakScott2021], we compared two types of models: *group models* and *individual models*. To assess the accuracy of body position classification for each recording session, we reserved the last 25% of a session's proximal comparison data as the testing set. The testing set was never used as training data. The two types of models differed based on what data were used during training, but the same random forest modeling procedure was used in both cases. A group model was trained for each session using an aggregated dataset containing the first 75% of all **other** sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set. In contrast, individual models were trained using the first 75% of a recording session, then tested against the last 25% of the same session. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches. Figure \@ref(fig:metrics), Table \@ref(tab:metrics-table), and Table \@ref(tab:metrics-by-class) summarize the performance of group and individual models using standard metrics for classification.

@@

Beyond accuracy, it is important to establish that classification is neither too sensitive nor too conservative in choosing one position versus another. Sensitivity refers to the proportion of events of a given position that were correctly identified (e.g., out of 100 human-coded sitting windows, how many of those windows did the model correctly classify as sitting?). High sensitivity means that events are unlikely to be missed. In contrast, positive predictive value (PPV) refers to the proportion of events classified for a given position that actually belonged to that position (e.g., if the model said a baby was upright during 100 windows, how many of those windows were indeed human-coded sitting events?). High PPV means that we can be confident in the event label. Table \@ref(tab:metrics-by-class) shows the sensitivity and PPV by body position class for group and individual models. For group models, sensitivity and PPV were similar: They were highest for supine, prone, and sitting (the most accurately identified class) and lowest for upright and held. All values were sufficient (> .8). Results for individual models were similar to group models, with the exception of a somewhat lower sensitivity score for held. Overall, the results suggest a reasonable balance between sensitivity and PPV among different body position classes for both model types. Table \@ref(tab:metrics-table) shows the overall sensitivity and PPV across classes.

@@

In the first set of results, we showed that group and individual models trained from data during the proximal comparison period were accurate at classifying body position in the same recording. In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal period (Figure \@ref(fig:exemplar-timeline)). A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used. As previously mentioned, distal videos had only coarse temporal synchrony with motion recordings (misalignment of up to ~ 5 s in either direction). This misalignment precluded calculating accuracy based on the proportion of matching events, so we used another technique following past work [@YaoPlotz2019; @FranchakScott2021]. For each session, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding. Below, we report correlations between the model-predicted and human-coded aggregated body position time across the entire distal period as well as within finer, 10-minute bins.

Whereas the proximal analyses used all `r length(unique(ppts$id_uni))` sessions, this was not possible in the distal comparison. Because the start of the visit was scheduled during a time the infant was awake, it was common for a nap to be needed following the proximal period. Nine sessions were excluded because the infant was either napping or otherwise not on camera during the entire 90-minute distal recording. Three additional sessions were excluded because a caregiver accidentally turned off the video camera (*n* = 1) or (purposefully) left the house (*n* = 2). This left `r length(unique(agree$group$file))` sessions with usable distal comparison data.

@@

Visual inspection of Figure \@ref(fig:part2overall) indicated two extreme outliers, which we marked by a gray square and a gray diamond. We closely investigated each of the two outliers to understand why their agreement was below the standard of the other sessions. The "gray square" outlier had significant confusion between sitting and supine classification. Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine. The infant also spent a long time in the mother's arms in an ambiguous supine/sitting position that was hard to classify. This participant's proximal accuracy was also poor because similar ambiguous appeared during the initial recording. In contrast, the "gray diamond outlier" had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories. Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as "upright" but the models predicted it as "held". Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infants moved while walking upright.

@@

```{r part2bins, warning=FALSE, fig.cap="Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison. Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins. Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute. Performance is plotted separately for (A) group models and (B) individual models. The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=8, fig.height=4.5, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
calculate_agreement_bins <- function(ds) {
  ds <- ds %>% group_by(file) %>% 
    mutate(sample = row_number(), bin = floor(sample/300)) %>% 
    add_count(file, bin) %>% ungroup() %>% filter(n > 210) %>% group_by(file) %>% mutate(num_bins = length(unique(bin))) %>% ungroup()
  model <- ds %>% count(file, bin, num_bins, pos) %>% add_count(file, bin,num_bins,  wt = n, name = "total") %>% mutate(prop_model = n/total)
  human <- ds %>% count(file, bin, num_bins,  code) %>% add_count(file, bin, num_bins, wt = n, name = "total") %>% mutate(prop_human = n/total)
  agree <- full_join(select(model, file, pos, prop_model, bin, num_bins), 
                     select(human, file, code, prop_human, bin, num_bins), 
                     by = c("file" = "file", "pos" = "code", "bin" = "bin", "num_bins" = "num_bins"))
  agree <- agree %>% mutate(prop_human = ifelse(is.na(prop_human), 0, prop_human),
                            prop_model = ifelse(is.na(prop_model), 0, prop_model),
                            pos - factor(pos, levels = pos_levels))
  agree <- complete(agree, nesting(file, bin, num_bins), pos, fill = list(prop_human = 0, prop_model = 0))
}

@@

Collecting real-time data that can be synchronized to other sources is a key goal in our development of long-form body position classification. Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale. We repeated the agreement analysis (correlating human-coded body position prevalence with model-predicted prevalence) after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)). Infants had varying numbers of 10-minute bins depending on much time they were awake and on camera. Bins were included only if there was > 7 minutes of usable data.

Table \@ref(tab:pt2binstable) shows the agreement correlation coefficients for group and individual models, including and excluding the two outliers identified in the previous section. Performance at a short timescale was similar to performance overall: Overall agreement after excluding outliers was excellent for both group (*r* = `r group_agree_cor_noout`) and individual models (*r* = `r split_agree_cor_noout`). Within-position correlations were weakest for held and strongest for upright regardless of the model type. Agreement for prone was better for group models, whereas sitting and supine were better predicted by individual models.

@@

Among participants who wore the sensors from morning to bedtime, the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \@ref(fig:timelines)). The total length of the recording period ranged from `r apa_num(min(coverage$total))`-`r apa_num(max(coverage$total))` hours with a median of `r apa_num(median(coverage$total))` hours. Nap times reported during the recording period ranged from `r apa_num(min(coverage$naptime))`-`r apa_num(max(coverage$naptime))` hours with a median of `r apa_num(median(coverage$naptime))` hours. Body position data was available to describe `r apa_num(min(coverage$prop_accounted)*100, digits = 1)`-`r apa_num(max(coverage$prop_accounted)*100, digits = 1)`% of the awake portion of the recording period (median = `r apa_num(median(coverage$prop_accounted)*100, digits = 1)`%). Younger infant \#13 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home. Overall, the recordings produced a median of `r apa_num(median(coverage$hours_recorded))` hours of motion data, with the entire dataset totaling `r apa_num(sum(coverage$hours_recorded))` hours. Even the shortest full recording (`r apa_num(min(coverage$hours_recorded))` hours) exceeded the longest observational sessions in past work in which an experimenter operated a camera.

@@

The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically. In @Survey, upright time was 0.6-5.5% of the time for infants 3-6 months, increasing to 22.0% at 12 months. Results from the current study were similar: younger infants' sitting time estimated from full-day group models was `r pos_sum_lookup["Upright", "Younger_Group"]`% compared with `r pos_sum_lookup["Upright", "Older_Group"]`% for the older group. In @Survey, prone time was 2.9-9.2% of the time for infants 3-6 months, and was 7.2% at 12 months. In the current study, prone time was `r pos_sum_lookup["Prone", "Younger_Group"]`% for younger infants and `r pos_sum_lookup["Prone", "Older_Group"]`% for older infants. Held, sitting, and supine are more difficult to compare because prior work [@Survey] considered held as any time that infants were in caregivers' arms or on their laps, however, we counted laying or sitting on caregivers' laps as supine and sitting, respectively. Further complicating the comparison, @Survey included a reclined category that we would classify as either supine or sitting (whichever it was closer to). Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants (Table \@ref(tab:agetable)). Sitting was greater for older infants (`r pos_sum_lookup["Sitting", "Older_Group"]`%) compared with younger infants (`r pos_sum_lookup["Sitting", "Younger_Group"]`%). Moreover, sitting was the most frequent body position followed by upright in 12-month-olds [@Survey], just as we observed in the current study. Overall, the estimates from the current investigation are consistent with age differences found using a different methodology.
